---
title: '{gt} package workshop'
author: "Juliette Verstaen"
date: "10/8/2020"
output: html_document
---

**Thanksgiving food survey data:**
https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-20
 
```{r setup, include=FALSE}
 
knitr::opts_chunk$set(echo = TRUE)
 
library(tidyverse) ## for sure need this one
library(stringr)
library(tm)
library(gt) ## for sure need this one
# library(webshot) ## this was me trouble shooting saving
# webshot::install_phantomjs() ## this was me trouble shooting saving
 
thanksgiving_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-20/thanksgiving_meals.csv')
 
```

Wrangle the data
There is a lot of naming mismatches unfortunatly, but we can fiz that!

Will make all cheese cake falvors cheese cakes, all custards custars
```{r}
thanksgiving_pie_int <- thanksgiving_raw %>% 
  dplyr::select(pie1:pie11, pie13) %>% 
  pivot_longer(cols =  pie1:pie13, names_to = "pies_num") %>% 
  select(type = value) %>% 
  table() %>%
  as_tibble() %>% 
  rename(Type = ".",
         Count = n) %>% 
  mutate(Type = tolower(Type)) %>% 
  mutate(Type = ifelse(Type == "blueberry or peach cobbler", "blueberry",
                       ifelse(Type == "cheesecak", "cheesecake",
                              ifelse(Type == "homemade pumpkin pie", "pumpkin", 
                                     ifelse(Type == "cherry cheescake", "cherry cheesecake", 
                                            ifelse(Type %in% c("lemon merang", "lemon mirangue"), "lemon meringue", 
                                                   ifelse(Type %in% c("mince meat", "minced meat"), "mincemeat", Type))))))) %>% 
  add_row(Type = "blueberry pie", Count = 1) %>% 
  add_row(Type = "banana cream pie", Count = 1) %>% 
  filter(Type != "blueberry pie, banana cream pie") %>% 
  mutate(Type = ifelse(str_detect(Type, "cheesecake"), "cheesecake", Type)) %>% 
  mutate(Type = removeWords(Type, "pie")) %>% 
  add_row(Type = "lemon meringue", Count = 1) %>% 
  add_row(Type = "mincemeat", Count = 1) %>% 
  filter(Type != "lemon maraigne, minch meat") %>% 
  mutate(Type = ifelse(str_detect(Type, "custard"), "custard", Type)) %>% 
  filter(Type != "usually some other  along with an apple ",
         Type != "we serve pine nut cake not ",
         Type != " with ice cream and whipped cream")%>% 
  add_row(Type = "pecan", Count = 1) %>% 
  add_row(Type = "mincemeat", Count = 1) %>% 
  filter(Type != "pecan, mincemeat") %>% 
  mutate(Type = str_trim(Type, side = c("both"))) %>% 
  group_by(Type) %>% 
  dplyr::summarise(Count = sum(Count)) %>% 
  ungroup() %>% 
  mutate(Type = ifelse(Count <5, "other", Type))%>% 
  group_by(Type) %>% 
  dplyr::summarise(Count = sum(Count)) %>% 
  ungroup() %>% 
  rename(Flavor = Type) %>% 
  mutate(Type = case_when(Flavor %in% c("buttermilk", "chocolate", "mincemeat", "other", "sweet potato") ~ "other",
                          Flavor == "none" ~ "none", 
                          T ~ "fruit"),
         r_ladies_count = Count) %>% 
  select(flavor = Flavor, type = Type, survey_counts = Count, r_ladies_count) 

thanksgiving_pie <- thanksgiving_pie_int %>% 
  mutate(survery_sum = sum(survey_counts),
         r_sum = sum(r_ladies_count)) %>% 
  rowwise() %>% 
  mutate(survey_percent = survey_counts/survery_sum,
          r_percent = r_ladies_count/r_sum)  %>% 
  select(-survery_sum, -r_sum)
  
```



R - Ladies {gt} workshop
 

Outline:
 
Intro:
- talk about all the different packages you can make tables in
- I'm not super familiar with any of them, I haven't had to make too many tables
- this workshop focus on {gt} package - grammer of tables (modeled after grammer of graphics ie: ggplot)
- show a bunch of pictures of plots that people made with gt
- show the parts of a gt table picture
 
chat about the data a bit and where it comes from. do more research on the transgender remeberence day meaning, history, etc


follow this https://towardsdatascience.com/exploring-the-gt-grammar-of-tables-package-in-r-7fff9d0b40cd and can also use this to help https://github.com/gkaramanis/tidytuesday/blob/master/2020-week40/beyonce-swift.R
 
this might be helpful too  https://evamaerey.github.io/data_manipulation/gt#37
 
-Talk about cols_move, cols_moce_to_end, cols_move_to_start
- cols_width
- color the rows/ columns
- cols_align
- summary_rows() might be useful. can do the total, or mean or whatever. can do it by one column and then the
- add music emoji icon in the title just for fun, or a record emoji, sing has a microphone https://github.com/hadley/emo
      - tab_header(title = md("&#127829; Pizzas Sold in 2015 &#127829;")
- data_color
- use a color pallett? - > paletteer is directly useable in this!
- save it! lots of ways to do so
 
 
 
# Step 1: Make a simple table (It really is that easy!)
```{r}

orange <- "#F7B060"
tan <- "#E2D0B8"
blue <- "#C1DDCE"


thanksgiving_pie %>% 

  ## chat about this after all the table formatting and getting into internal formatting, a good reminder that we are using a df to creat the table, can modify !
  
  dplyr::arrange(desc(survey_percent)) %>% 
 # group_by(type) %>% 
  
  ## make the table!
  gt() %>% 
  
  ## add in a title and subtitle
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  
  ## add in a source for the entire table
  tab_source_note(source_note = "Source: fivethirtyeight survery from 2015") %>% 
  
  ## or we can specify a column source as an actual footnote
  tab_footnote(footnote = ("Source: fivethirtyeight"),
    locations = cells_column_labels(columns = vars(survey_counts))) %>% 

  ## move columns around
  cols_move(columns = "survey_percent", after = "survey_counts") %>% 
    
  ## name all your columns
  cols_label(survey_counts = "Source Counts", 
             r_ladies_count = "R-Ladies Counts",
             survey_percent = "Source Percent",
             r_percent = "R-Ladies Percent",
             flavor = "Pie Flavor",
             type = "Pie Type")  %>% 
  
  ## add in banners
  tab_spanner(label = "fivethirtyeight",
              columns = vars(survey_counts, survey_percent)) %>%  ### WEIRD, you have to use the OG data name, not the renamed column label
  tab_spanner(label = "R-Ladies SB",
              columns = vars(r_ladies_count, r_percent)) %>% 
  
  ## format the percent columns. there are formatting functions for date/time, currency, scientific notation, numbers, etc. we are going to format these percentages
  fmt_percent(columns = vars(survey_percent, r_percent),
              decimals = 2) %>% 
  
  ## we can use colors to highlight what we want the viewer to keep with them as a main take away
  ## can color variable columns based on the data reprsented 
    data_color(columns = vars(r_ladies_count, survey_counts), ## can do this for 1 or more columns
    colors = scales::col_numeric(palette = c(blue, tan, orange), # A color scheme (gradient)
       domain = c(0, 740))) %>%  # Column scale endpoints
  
:summary_rows(columns = vars(survey_counts, r_ladies_count),
             fns = list(Total = "sum"),
             formatter = fmt_number())
            
            #missing_text = "---",)


## can also color a certain cell based on criteria 

## or you can highlight an entire row based on a column value
  ## color in the fruit
  # tab_style(
  #   style = cell_fill(color = orange),
  #   locations = cells_body(
  #     rows = type == "fruit")) 
  # 
  
summary_rows(groups = FALSE,
             columns = vars(survey_counts, r_ladies_count),
             fns = list(Total = "sum"))
```
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 ## for saving the file i cant do it in mazu for some strange reason
# test <- expeditions_raw %>%
#   dplyr::select(peak_name, year, season) %>%
#   head(20) %>%
#   gt()
#
# gtsave(test, "test.pdf")
# gtsave(test, "test.png")
