---
title: '{gt} package workshop'
author: "Juliette Verstaen"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 

 
 https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-20
 
 
```{r setup, include=FALSE}
 
knitr::opts_chunk$set(echo = TRUE)
 
library(tidyverse)
library(dplyr)
library(stringr)
library(gt)
library(webshot)
webshot::install_phantomjs()
 
thanksgiving_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/blob/master/data/2018/2018-11-20/thanksgiving_meals.csv')
 
```

R - Ladies {gt} workshop
 

Outline:
 
Intro:
- talk about all the different packages you can make tables in
- I'm not super familiar with any of them, I haven't had to make too many tables
- this workshop focus on {gt} package - grammer of tables (modeled after grammer of graphics ie: ggplot)
- show a bunch of pictures of plots that people made with gt
- show the parts of a gt table picture
 
1) make a table with just the album and the release date - focus only on beyonce
  - the release date can we tidied directly in gt table
    -  fmt_date( for date formating), can also do fmt_currency, fmt_time, fmt_number etc
2) next we want to add a title, maybe a subtitle for shits and giggles, and maybe also show how to add a caption
3) next we want to add the format types the album was released in
4) add in taylor swift so it looks semi faceted/ like two tables
 
follow this https://towardsdatascience.com/exploring-the-gt-grammar-of-tables-package-in-r-7fff9d0b40cd and can also use this to help https://github.com/gkaramanis/tidytuesday/blob/master/2020-week40/beyonce-swift.R
 
this might be helpful too  https://evamaerey.github.io/data_manipulation/gt#37
 
-Talk about cols_move, cols_moce_to_end, cols_move_to_start
- cols_width
- color the rows/ columns
- cols_align
- summary_rows() might be useful. can do the total, or mean or whatever. can do it by one column and then the
- add music emoji icon in the title just for fun, or a record emoji, sing has a microphone https://github.com/hadley/emo
      - tab_header(title = md("&#127829; Pizzas Sold in 2015 &#127829;")
- data_color
- use a color pallett? - > paletteer is directly useable in this!
- save it! lots of ways to do so
 
 
 
 
 
 
 
# test <- expeditions_raw %>%
#   dplyr::select(peak_name, year, season) %>%
#   head(20) %>%
#   gt()
#
# gtsave(test, "test.pdf")
# gtsave(test, "test.png")
