---
title: '{gt} package workshop'
author: "Juliette Verstaen"
date: "Novemeber 18, 2020"
output: html_document
---

# About: 

**Welcome!**

In this workshop we will go over the basics of the {gt} package using FiveThirtyEight Thanksgiving Dinner or Transgender Day of Remembrance (TDoR) survey [data](https://github.com/rfordatascience/tidytuesday/tree/master/data/2018/2018-11-20). This workshop is just a sample of the things that can be done with the {gt} package. For a full list of functions and how to use them I recommend this [R Studio info webpage](https://gt.rstudio.com/index.html), which is surprisingly helpful. The 'Reference' tab has a full list of functions.

Any questions, comments, or concerns please feel free to contact me at verstaen@nceas.ucsb.edu.

## Set up: libraries and read in data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 
## libraries


## read in the wrangled data 
 

```


## Step 1: Turn the df into a table with just one line!
```{r basic table}



```

## Step 2. Add in a title and subtitle
```{r titles and subtitles}

## add a title and subtitle  
thanksgiving_pie %>% 
  gt() %>% 

  

## we can bold or italicize directly into the titles using md(). one "*" sandwiching a word or phrase will italicize, and two "**" will bold
thanksgiving_pie %>% 
  gt() %>% 
  

 
```

## Step 3: Captions and footnotes
```{r captions and footnotes}

## create captions/notes for entire table
thanksgiving_pie %>% 
  gt() 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
    

## maybe we want to pull out specific cells, or columns with a footnote  
thanksgiving_pie %>% 
  gt() %>% 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
    
```

## Step 4. Rearrange columns, create column banners, rename columns, and align

There is much more you can do in regards to formatting the table/ columns to fit your needs. I highly recommend checking out the [R Studio info webpage](https://gt.rstudio.com/index.html) to see what options are available to you.

```{r column formating}

thanksgiving_pie %>% 
  gt() %>% 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  tab_source_note(source_note = "Table created 11/18/20 for R-Ladies SB") %>% 
  tab_footnote(footnote = ("Source: FiveThirtyEight"),
    locations = cells_column_labels(columns = vars(survey_counts, survey_percent))) %>% 
  
  ## move columns around 
  
  ## add in banners
  
  ## rename columns
  
  ## align columns
  
  
```
*NOTE*: Even once you have changed the column labels, that column name in the df is still the original name. So, when using functions that refer to the columns moving forward, we will still need to refer to the original column labels. This is kind of nice because it gives you freedom in regards to labeling.


## Step 5: Formating data in columns
There are a lot of data formatting functions to apply to columns in the {gt} package. You can directly format percentages, dates, time, currency, scientific notation, numbers, etc. All these functions start with "fmt_". 

```{r data formating}

thanksgiving_pie %>% 
  gt() %>% 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  tab_source_note(source_note = "Table created 11/18/20 for R-Ladies SB") %>% 
  tab_footnote(footnote = ("Source: FiveThirtyEight"),
    locations = cells_column_labels(columns = vars(survey_counts, survey_percent))) %>% 
  cols_move(columns = "survey_percent", after = "survey_counts") %>% 
  tab_spanner(label = "FiveThirtyEight",
              columns = vars(survey_counts, survey_percent)) %>% 
  tab_spanner(label = "R-Ladies SB",
              columns = vars(r_ladies_counts, r_percent)) %>% 
  cols_label(survey_counts = "Counts", 
             r_ladies_counts = "Counts",
             survey_percent = "Percent",
             r_percent = "Percent",
             flavor = "Flavor",
             type = "Type")  %>% 
  cols_align("center") %>% 
  
  ## format the percent

  
```

Lets organize the rows in descending order for survey_percentage. Remember, to create a gt table we start with a regular data frame that can be wrangled anyway we want before using gt() to convert it into a table. Here we will arrange the df before using gt()

```{r arranging row order}

thanksgiving_pie %>% 
  
  ## arrange row order


  gt() %>% 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  tab_source_note(source_note = "Table created 11/18/20 for R-Ladies SB") %>% 
  tab_footnote(footnote = ("Source: FiveThirtyEight"),
    locations = cells_column_labels(columns = vars(survey_counts, survey_percent))) %>% 
  cols_move(columns = "survey_percent", after = "survey_counts") %>% 
  tab_spanner(label = "FiveThirtyEight",
              columns = vars(survey_counts, survey_percent)) %>% 
  tab_spanner(label = "R-Ladies SB",
              columns = vars(r_ladies_counts, r_percent)) %>% 
  cols_label(survey_counts = "Counts", 
             r_ladies_counts = "Counts",
             survey_percent = "Percent",
             r_percent = "Percent",
             flavor = "Flavor",
             type = "Type")  %>% 
  cols_align("center") %>% 
  fmt_percent(columns = vars(survey_percent, r_percent),
              decimals = 2) 

```


## Step 6: Using colors in your table

These are the colors I know I'll want to use; you can also use ggplot colors or palleteer pallets directly! Amazing!
```{r specify colors}
orange <- "#F7B060"
tan <- "#E2D0B8"
blue <- "#C1DDCE"
```

Lets say we really want to highlight the differences in counts for each pie flavor. 
```{r apply colors}

thanksgiving_pie %>% 
  dplyr::arrange(desc(survey_percent)) %>%  
  gt() %>% 
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  tab_source_note(source_note = "Table created 11/18/20 for R-Ladies SB") %>% 
  tab_footnote(footnote = ("Source: FiveThirtyEight"),
    locations = cells_column_labels(columns = vars(survey_counts, survey_percent))) %>% 
  cols_move(columns = "survey_percent", after = "survey_counts") %>% 
  tab_spanner(label = "FiveThirtyEight",
              columns = vars(survey_counts, survey_percent)) %>% 
  tab_spanner(label = "R-Ladies SB",
              columns = vars(r_ladies_counts, r_percent)) %>% 
  cols_label(survey_counts = "Counts", 
             r_ladies_counts = "Counts",
             survey_percent = "Percent",
             r_percent = "Percent",
             flavor = "Flavor",
             type = "Type")  %>% 
  cols_align("center") %>% 
  fmt_percent(columns = vars(survey_percent, r_percent),
              decimals = 2) %>% 
  
  ## highlighting differences in the quantitative data 
 
  
  ## what if we wanted to highlight the pies that are fruit types

  
```


## Step 7: Grouping and summary stats

Grouping can be done a couple ways in a gt table object: tab_row_group(), group_by() in dplyr, or specifying group in the gt() call

```{r grouping and faceting}

thanksgiving_pie %>% 
  dplyr::arrange(desc(survey_percent)) %>%  
  
  ## specify grouping variable and column

  gt(    ) %>% 
  
  tab_header(title = "What pies do we eat at Thanksgiving?",
             subtitle = md("*may the slice be with you*")) %>% 
  tab_source_note(source_note = "Table created 11/18/20 for R-Ladies SB") %>% 
  tab_footnote(footnote = ("Source: FiveThirtyEight"),
    locations = cells_column_labels(columns = vars(survey_counts, survey_percent))) %>% 
  cols_move(columns = "survey_percent", after = "survey_counts") %>% 
  tab_spanner(label = "FiveThirtyEight",
              columns = vars(survey_counts, survey_percent)) %>% 
  tab_spanner(label = "R-Ladies SB",
              columns = vars(r_ladies_counts, r_percent)) %>% 
  cols_label(survey_counts = "Counts", 
             r_ladies_counts = "Counts",
             survey_percent = "Percent",
             r_percent = "Percent",
             flavor = "Flavor",
             type = "Type")  %>% 
  cols_align("center") %>%
  fmt_percent(columns = vars(survey_percent, r_percent),
              decimals = 2) %>% 

  data_color(columns = vars(r_ladies_counts, survey_counts),
              colors = scales::col_numeric(palette = c(blue, tan, orange),
                                           domain = c(5, 730))) %>% 
  
    ## calculate summary statistics



```


# Other cool stuff:

* giving the stub head a label (tab_stubhead()), and relabeling row names and such 

* You can use emojis in your tables!! It's super easy https://github.com/hadley/emo

* {palleteer} package can be used directly in the gt function to grab awesome colors

* You can save your creation with great resolution using gtsave! For some reason figuring this out involved weird trouble shooting for me and I had to download {webshot} and then run this line once [webshot::install_phantomjs()] before I was able to save.

